Self recovering capabilities in images using the BPCS Steganography algorithm

### 2014-10-07 Paper/main.tex
Completed most of ch2 changes only def's missing and one thing to move

\subsection{Introduction}
//TODO
Self-recovering images are images that have their own image content embedded into themselves using image steganography techniques. 
This content can then be extracted from the image at a later stage in order to reconstruct the image to an approximate original state in the case that the image was damaged.

\hspace{0pt} \\
In order to understand the self-recovering system proposed in this dissertation, one has to understand the underlying technologies.  This chapter will look at image steganography in more detail in section \ref{introSteganography} and at watermarking in section \ref{introFragWatermarking}  Self-embedding and recovery will be discussed in section \ref{introSelfEmbedRecovery} as well as the inherent quality trade off in section \ref{introQualityTrade} with a conclusion in section \ref{introConclusion}

Self recovering images poses an interesting problem.
\hspace{0pt} \\
//TODO insert formal definition here
\hspace{0pt} \\
There is an inherent trade-off between the quality of the image after embedding and the quality of the image after recovery.

\iffalse
//TODO possibly remove this
This report proposes a method in which a high capacity image steganography embedding algorithm (described in \ref{introHighCap}) is applied to the problem of self-recovering images to attempt to increase the quality of the images (described in \ref{introQualityIncrease}) both after embedding and after recovery.
\fi

@@

\subsection{Fragile Watermarking}
\label{introFragWatermarking}
Fragile watermarking is a variation of traditional watermarking where a watermark is embedded into an image such that subsequent alterations to the watermarked image can be detected with high probability \cite{lin1999review}.
This means that if modifications are made to the watermarked image it would be possible to detect that modifications were made as well as what parts of the image were modified.
This is a necessary component in self recovering images because in order to be able to recover image data it first needs to be determined if the image after embedding has been modified and if so what parts of the image was modified.
The proposed method will use this information to determine which parts of the image should be recovered from the image reference content.
A fragile watermark is used for this purpose.

@@

\subsection{Quality Trade-off}
\label{introQualityTrade}
There is an inherent quality trade-off present within self embedding schemes.
The trade-off exists between the image after embedding and the image after reconstruction.
If an embedding algorith such as LSB(Least significan bit) embedding is used (as in \ref{ErasureChannelOverview}) the actual bit values of the original image are overwritten.
Naturally the more bit values are overwritten the more noise is introduced into the image.
This effectively reduces the quality of the image after embedding.

@@

\subsubsection{Increasing quality with capacity}
\label{introQualityIncrease}
It might be possible to increase the quality of the image after reconstruction while still keeping the quality after embedding high.
It is proposed that applying an embedding algorithm designed for high capacity while retaining quality (as discussed in \ref{introHighCap}) to the problem of self recovering could possibly increase the quality of the image after embedding as well as after reconstruction.
This is because such an algorithm should provide higher embedding capacity while producing less noise in the image.

\subsubsection{High capacity embedding algorithms}
\label{introHighCap}
The proposed method in this paper attempts to apply the algorithm for BPCS Steganography (Bit-Plane Complexity Segmentation Steganography) \cite{beaullieubpcs} to the problem of self recovering images, specifically to increase overall quality.

\hspace{0pt} \\
The goal of the BPCS algorithm is to embed as much data as possible into a cover image without detection \cite{beaullieubpcs}.
This refers to detection by human perception as well as statistical analysis although the latter is not the focus of this paper.
The BPCS algorithm uses an adaptive scheme to classify blocks that are suitable for embedding (similar to \ref{DynamicBlockOverview}).
BPCS steg differs however in the fact that the blocks can be conjugated in a checker board pattern to increase complexity and decrease detectability if needs be.
BPCS steg also differs in the way in which embedding takes place.
Instead of always embedding into the 3-LSB planes as in \ref{DynamicBlockOverview}.
Bit planes are dynamically allocated for embedding.
One section may thus only embed into one bit plane while another may use all 7 available bit planes.

\hspace{0pt} \\
The BPCS algorithm can achieve a very high capacity while still remaining undetectable through human perception.
This makes the BPCS algorithm a possible good choice to increase the capacity of a self recovering image scheme and possible increase the quality through this process.

\subsection{Conclusion}
\label{introConclusion}
//TODO

@@

\subsubsection{BPCS Steg}
\label{bpcsSteg}
In order to be able to reconstruct the image after it has been tampered with the image needs to be embedded into itself. 
For this embedding the proposed method used BPCS Steganography \cite{beaullieubpcs}.

\hspace{0pt} \\
This method works with the following steps:
\begin{enumerate}
  \item Get the binary representation of the image
  \item Convert the binary representation into Gray code
  \item Divide the image into bit planes
  \item Divide the bit planes into 8x8 pixel blocks
  \item Calculate the complexity for each block
  \item Divide the image reference content to be embedded into similar binary 8x8 blocks
  \item Calculate the complexity for each image reference block and conjugate if not complex enough
  \item for each image block that is complex enough, replace that block with the next image reference content block
  \item Convert the blocks back to binary representation
  \item reconstruct the image from the processed blocks 
\end{enumerate}
The proposed method works with colour images which means that for each pixel there is a red, green and blue value. This means that each block effectively stores 3 bits for each pixel.

\hspace{0pt} \\
BPCS Steg has an inherent property that if any part of the image is modified the embedded message will be lost.
This would obviously be a problem for the proposed method since the image will definitely be modified.

>>>>

\subsection{Introduction}
Self-recovering images are images that have their own image content embedded into themselves using image steganography techniques. 
This content can then be extracted from the image at a later stage in order to reconstruct the image to an approximate original state in the case that the image was damaged.

\hspace{0pt} \\
In order to understand the self-recovering system proposed in this dissertation, one has to understand the underlying technologies.  This chapter will look at image steganography in more detail in section \ref{introSteganography} and at watermarking in section \ref{introFragWatermarking}. Self-embedding and recovery will be discussed in section \ref{introSelfEmbedRecovery} with a conclusion in section \ref{introConclusion}

Self recovering images poses an interesting problem.
\hspace{0pt} \\
//TODO <add formal definition here>
\hspace{0pt} \\
There is an inherent trade-off between the quality of the image after embedding and the quality of the image after recovery.

@@

\hspace{0pt} \\
The proposed method uses steganography to embed the information into the image in a way that is not easily visibly detectable by a human observer.
A successful stegnographic method has certain characteristics:
\begin{itemize}
	\item Invisibility: How easy it is to detect that information has been embedded using steganography.
	\item Robustness: How much of the image that is embedded into can be modified without compromising the embedded information.
	\item Capacity: The maximal number of bits that can be embedded without introducing statistically detectable artifacts \cite{fridrich2009steganography}. It should also be added that the embedding must also not be visually detectable.
\end{itemize}
Least significant bit embedding (LSB) is a stegnographic method that can be used to embed information. 
LSB works by replacing the least significant bit (or n least significant bits) of an image colour value with the binary bits from a message. The least significant bit of a colour value is the binary bit/s that carry the least value.

\hspace{0pt} \\
BPCS Steganography is another stegnographic embedding method \cite{beaullieubpcs}.

\hspace{0pt} \\
This method works with the following steps:
\begin{enumerate}
  \item Get the binary representation of the image
  \item Convert the binary representation into Gray code
  \item Divide the image into bit planes
  \item Divide the bit planes into 8x8 pixel blocks
  \item Calculate the complexity for each block
  \item Divide the image reference content to be embedded into similar binary 8x8 blocks
  \item Calculate the complexity for each image reference block and conjugate if not complex enough
  \item for each image block that is complex enough, replace that block with the next image reference content block
  \item Convert the blocks back to binary representation
  \item reconstruct the image from the processed blocks 
\end{enumerate}

\hspace{0pt} \\
The BPCS algorithm is a high capacity embedding algorithm. The goal of the BPCS algorithm is thus to embed as much data as possible into a cover image without detection \cite{beaullieubpcs}.
This refers to detection by human perception as well as statistical analysis although the latter is not the focus of this paper.
The BPCS algorithm uses an adaptive scheme to classify blocks that are suitable for embedding (similar to \ref{DynamicBlockOverview}).
BPCS steg differs however in the fact that the blocks can be conjugated in a checker board pattern to increase complexity and decrease detectability if needs be.
BPCS steg also differs in the way in which embedding takes place.
Instead of always embedding into the 3-LSB planes as in \ref{DynamicBlockOverview}.
Bit planes are dynamically allocated for embedding.
One section may thus only embed into one bit plane while another may use all 7 available bit planes.

\subsection{Fragile Watermarking}
//TODO <add formal definition here> 
\label{introFragWatermarking}
Fragile watermarking is a variation of traditional watermarking where a watermark is embedded into an image such that subsequent alterations to the watermarked image can be detected with high probability \cite{lin1999review}.
This means that if modifications are made to the watermarked image it would be possible to detect that modifications were made as well as what parts of the image were modified.
This is a necessary component in self recovering images because in order to be able to recover image data it first needs to be determined if the image after embedding has been modified and if so what parts of the image was modified.

\iffalse 
	//TODO this needs to move to a different section
	The proposed method will use this information to determine which parts of the image should be recovered from the image reference content.
	A fragile watermark is used for this purpose. 
\fi

@@

\hspace{0pt} \\
There is an inherent quality trade-off present within self embedding schemes.
The trade-off exists between the image after embedding and the image after reconstruction.
If an embedding algorith such as LSB(Least significan bit) embedding is used the actual bit values of the original image are overwritten.
Naturally the more bit values are overwritten the more noise is introduced into the image.
This effectively reduces the quality of the image after embedding.

@@

\hspace{0pt} \\
It might be possible to increase the quality of the image after reconstruction while still keeping the quality after embedding high.
By using a high capacity embedding algorithm the overall embedding capacity is increased and a higher quality restoration image can be embedded.

\hspace{0pt} \\
The proposed method in this paper attempts to apply the algorithm for BPCS Steganography
to the problem of self recovering images, specifically to increase overall quality.
The BPCS algorithm can achieve a very high capacity while still remaining undetectable through human perception.
This makes the BPCS algorithm a good choice to increase the capacity of a self recovering image scheme and possibly increase the quality through this process.

\subsection{Conclusion}
\label{introConclusion}
This chapter gave important background information regarding staganography in general as well as two stegnographic embedding methods, LSB embedding and BPCS steganography in section \ref{introSteganography}.
Fragile watermarking was discussed in section \ref{introFragWatermarking} as well as self embedding and recovery in section \ref{introSelfEmbedRecovery}.
The next chapter will discuss related work in existing self embedding techniques.

@@

The proposed method works with colour images which means that for each pixel there is a red, green and blue value. This means that each block effectively stores 3 bits for each pixel.

\subsubsection{BPCS steganography} 
\label{bpcsSteg}

BPCS Steg has an inherent property that if any part of the image is modified the embedded message will be lost.
This would obviously be a problem for the proposed method since the image will definitely be modified.
